Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-10-04 arxiv/proof_examples.tex
revise

 Now we expand the first term on the right-hand side of \eqref{eq_mtl_A12}: 
 \begin{align*}
 & \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)+ \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} + A X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
 &=h_1(A) + h_2(A) +h_3(A) + 2h_4(A) + 2h_5(A) + 2h_6(A) ,
 \end{align*}
 where
 \begin{align*}
&h_1(A):=  (\beta_1-A\beta_2)^\top (X^{(2)})^\top X^{(2)}  \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)  ,\\
& h_2(A) := {(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon^{(1)}}  ,\\
& h_3(A):=  A^2 { (\epsilon^{(2)} )^\top X^{(2)} \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)}} , \\
& h_4(A):=  (\epsilon^{(1)})^\top\left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right) X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1), \\
&h_5(A):=A (\epsilon^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1),\\
&h_6(A):=A (\epsilon^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top\left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)}.
\end{align*}
Next we estimate each term using Lemma \ref{largedeviation} in Appendix \ref{app_tool}.

@@

%control the difference between $g(A)$ and its partial expectation. 
 \begin{claim}\label{claim_largedev2}
 In the setting of Proposition \ref{cor_MTL_loss}, if $\hat A=\OO(1)$, then we have that with high probability,
$$\left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat A)\right| \le p^{-1/2+e}\cdot p\left(\sigma^2 +\kappa^2 +d^2 \right),$$
for any small constant $e>0$. Here we introduced the function
\begin{align*}
	\cal L(\hat A)	:=  \left[(\hat A-1)^2 \kappa^2 + (\hat A^2+1)\frac{d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} } +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).
\end{align*}
 \end{claim}
 \begin{proof}
 The proof of this claim is very similar to the one for Claim \ref{claim_largedev1}, where the only difference is that $\hat A$ now may depend on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. In this case, we can still use  Lemma \ref{largedeviation} to conclude the proof because $\hat A$ can be pulled out as a coefficient. Here we only give an outline of the proof without writing down all the details. We expand $L(\hat{\beta}_2^{\MTL}) $ in \eqref{Lbeta_HPS} as:
 \begin{align*}
L(\hat{\beta}_2^{\MTL}) &\cor  =  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat A\beta_1-\hat A^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat A   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2\\
&=h_1(A) + h_2(A) +h_3(A) + 2h_4(A) + 2h_5(A) + 2h_6(A) ,
\end{align*}
where

>>>>

 Now we expand the first term on the right-hand side of \eqref{eq_mtl_A12}: 
 \begin{align}
 & \left\| X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)+ \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} + A X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
 &=h_1(A) + h_2(A) +h_3(A) + 2h_4(A) + 2h_5(A) + 2h_6(A) ,\label{expand_6}
 \end{align}
 where
 \begin{align*}
&h_1(A):=  (\beta_1-A\beta_2)^\top (X^{(2)})^\top X^{(2)}  \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)  ,\\
& h_2(A) := {(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon^{(1)}}  ,\\
& h_3(A):=  A^2 { (\epsilon^{(2)} )^\top X^{(2)} \Sigma(A)^{-1}  (X^{(1)})^\top X^{(1)}\Sigma(A)^{-1} (X^{(2)})^\top \epsilon^{(2)}} , \\
& h_4(A):=  (\epsilon^{(1)})^\top\left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right) X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1), \\
&h_5(A):=A (\epsilon^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1),\\
&h_6(A):=A (\epsilon^{(2)})^\top X^{(2)}\Sigma(A)^{-1} (X^{(1)})^\top\left(A^2 X^{(1)}\Sigma(A)^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)}.
\end{align*}
Next we estimate each term using Lemma \ref{largedeviation} in Appendix \ref{app_tool}.

@@

%control the difference between $g(A)$ and its partial expectation. 
 \begin{claim}\label{claim_largedev2}
 In the setting of Proposition \ref{cor_MTL_loss}, if $\hat A=\OO(1)$, then we have that with high probability,
$$\left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat A)\right| \le p^{-1/2+e} \left(\sigma^2 +\kappa^2 +d^2 \right),$$
for any small constant $e>0$. Here we introduced the function
\begin{align*}
	\cal L(\hat A)	:=  \left[(\hat A-1)^2 \kappa^2 + (\hat A^2+1)\frac{d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} } +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).
\end{align*}
 \end{claim}
 \begin{proof}
 The proof of this claim is very similar to the one for Claim \ref{claim_largedev1}, where the only difference is that $\hat A$ now may depend on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. In this case, we can still use  Lemma \ref{largedeviation} to conclude the proof because $\hat A$ can be pulled out as a coefficient. Here we only give an outline of the proof without writing down all the details. Recall that $\beta_0$ is the shared component of $\beta_1$ and $\beta_2$ with i.i.d. Gaussian entries of mean zero and variance $p^{-1}\kappa^2$. Moreover, we denote the task-specific components by $\wt\beta_1$ and $\wt\beta_2$, whose entries i.i.d. Gaussian random variables of mean zero and variance $p^{-1} d^2 / 2$. Then we write $L(\hat{\beta}_2^{\MTL}) $ in \eqref{Lbeta_HPS} as:
 \begin{align}
L(\hat{\beta}_2^{\MTL})  =&  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat A -\hat A^2)\beta_0+(X^{(1)})^\top X^{(1)} \hat A\wt \beta_1 - (X^{(1)})^\top X^{(1)}  \hat A^2\wt \beta_2 \right] \right. \\
&\left. + (\Sigma^{(2)})^{1/2}\hat \Sigma^{-1}\left[ (X^{(2)})^\top \epsilon^{(2)} + \hat A   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2. \label{expand_15}
\end{align}
As in \eqref{expand_6}, we can expand this expression into the sum of 15 terms, and bound each term as in \eqref{eq_hA1}-\eqref{eq_hA56}. For example, for the main term $\hat A^2 \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1$, using 
Lemma \ref{largedeviation} and \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability for any small constant $e>0$:
\begin{align*}
&\hat A^2 \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1 - d^2 \cdot p^{-1}\bigtr{(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}}\right| \\
&\le p^e\cdot p^{-1}d^2 \cdot \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)} \right\|_F \\
&\le p^{-1/2+\e}d^2 \cdot  \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)} \right\|  \lesssim p^{-1/2+\e}d^2.
\end{align*} 
For the cross term $\hat A \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \epsilon^{(2)}$, using Lemma \ref{largedeviation} and \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability for any small constant $e>0$:
\begin{align*}
&|\hat A |\cdot \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \epsilon^{(2)}\right| \le p^e \cdot \sigma\sqrt{p^{-1}d^2} \cdot \|(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \|_F \\
&\lesssim p^{e}\sigma d \cdot  \|(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \| \lesssim p^{-1/2+e}\sigma d\le p^{-1/2+e}(\sigma^2+ d^2).
\end{align*}
The rest of the terms in the expansion of \eqref{expand_15} can be bounded in the same way, and we omit the details.
 \end{proof}
